Working with table environments in texor

### 2023-09-06 supplement/math-env/math-example.tex
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

The core goal of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. The key design goals are \citep{mathjax}:
\begin{itemize}
    \item High-quality display of mathematics notation in all browsers.
    \item No special browser setup required.
    \item Support for LaTeX, MathML, and other equation markup directly in the HTML source.
    \item An extensible, modular design with a rich API for easy integration into web applications.
    \item Support for accessibility, copy and paste, and other rich functionality.
    \item Interoperability with other applications and math-aware search.
    \item Support for equation conversion outside a browser (e.g., preprocessing on a server).
\end{itemize}

\subsubsection{Pandoc handling, extensions}
As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the R Markdown file into HTML,
the \verb|rjtools::rjournal_web_article| template uses MathJax as the math engine by default. 
We also specify the Mathjax version in the metadata of the generated Rmarkdown file.

\section{Inline math}
One can define inline math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
It is also possible to use inline math within captions and  
\begin{table}[htbp]
  \centering
  \begin{tabular}{| c | c |}
  \hline
  Command & Math\\
    \hline
    \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
\(\mu = (0,0,0)^\top \)
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    \( \mu = (0,0,0)^\top \)
    \end{minipage}\\
    \hline

@@

\begin{verbatim}
\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:1}
\end{align}
\end{verbatim}
The code above will render as Figure~\ref{eq:1}.

@@

Equation labels must start with the prefix \verb|eq:| in bookdown. All labels in bookdown must only contain alphanumeric characters, \verb|:|, \verb|-|, and/or \verb|/| as suggested in \citep{bookdown}.

@@

%Also avoid custom math commands MathJaX will never be able to parse it. The reason being
%MathJax renders math case by case and custom commands will already be removed by pandoc
%during the conversion, hence MathJax will not be aware of such redefinations.
To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
However, this will not work if you use functions/environments from CTAN packages or 
the ones that are out-of-the-box functions which do not render in HTML/markdown.

>>>>

The core goal of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. 
The key advantages are \citep{mathjax}:
\begin{itemize}
    \item High-quality display of mathematics notation in all browsers.
    \item No special browser setup required.
    \item Support for LaTeX, MathML, and other equation markup directly in the HTML source.
    \item Support for accessibility, copy and paste, and other rich functionality.
    \item Interoperability with other applications and math-aware search.
\end{itemize}

As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the R Markdown file into HTML,
the \verb|rjtools::rjournal_web_article| template uses MathJax as the math engine by default. 
We also specify the Mathjax version in the metadata of the generated Rmarkdown file.

\section{Inline math}
One can define inline math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
\begin{table}[htbp]
  \centering
  \begin{tabular}{| c | c |}
  \hline
  Command & Math\\
    \hline
    \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
\(\mu = (0,0,0)^\top \)
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    \( \mu = (0,0,0)^\top \)
    \end{minipage}\\
    \hline

@@

\begin{verbatim}
\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:1}
\end{align}
\end{verbatim}
The code above will render as Equation~\ref{eq:1}.

@@

Equation labels must start with the prefix \verb|eq:| in bookdown. All labels in bookdown must only contain alphanumeric characters, \verb|:|, \verb|-|, and/or \verb|/| as suggested in \citep{bookdown}. To accomadate this, the \pkg{texor} package has a lua filter implementation to correct and modify equation labels and references to the
bookdown accepted format. For example lets say this equation has a non standard label
\paragraph{LaTeX source code}
\begin{example}
\begin{equation}\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}

\end{equation}
The reference to above equation \eqref{binomial pdf}
\end{example}

\paragraph{Modified R markdown using \pkg{texor}}
\begin{verbatim}
$$\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}   (\#eq:binomial-pdf)$$

The reference to above equation \@ref(eq:binomial-pdf)
\end{verbatim}

\paragraph{Rendering of the above example}
\begin{equation}\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
\end{equation}

The reference to above equation \eqref{binomial pdf}

@@

The existence of user-defined LaTeX commands intended for the Math environments will work as long as they do not contain non-math LaTeX commands or commands from other CTAN math packages.

### 2023-09-06 supplement/metadata/metadata.tex
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

\section{Miscellaneous}

@@

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\href{www.google.com}{Google}| & \href{www.google.com}{Google} \\ \hline
\verb|\url{www.google.com}| & \url{www.google.com}  \\ \hline
\verb|Table \ref{table:3}| & Table \ref{table:3}  \\ \hline
\end{tabular}
\caption{Links}
\label{table:4}
\end{table}

\subsection{Colored text}
Some commands to describe colored text also work during conversion, such as
\verb|\textcolor{blue}{Blue Text}|  \textcolor{blue}{Blue Text}.

>>>>

\section{Miscellaneous}\label{misc}

@@

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\href{www.google.com}{Google}| & \href{www.google.com}{Google} \\ \hline
\verb|\url{www.google.com}| & \url{www.google.com}  \\ \hline
\verb|Table \ref{table:3}| & Table \ref{table:3}  \\ \hline
\verb|Section \nameref{misc}| & Section \nameref{misc}\\ \hline
\verb|\autoref{table:3}| & \autoref{table:3} \\ \hline
\end{tabular}
\caption{Links}
\label{table:4}
\end{table}

\paragraph{Note}
Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| will not work as intended in HTML.

\subsubsection{Lists}

An example of usage of lists in LaTeX and \pkg{texor} generated HTML.

\paragraph{itemize}
\begin{itemize}
\item A
\item B
\end{itemize}
\paragraph{enumerate}
\begin{enumerate}
\item A
\item B
\end{enumerate}
\paragraph{description}
\begin{description}
\item[\pkg{rebib}] Convert and Aggregate Bibliographies
\item[\pkg{texor}] Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'
\end{description}

\paragraph{Note:}
Some options for lists such as special numbering schemes or markups to description
item points might not work as intended and the generated HTML might be simple.

\subsubsection{Colored text}
Some commands to describe colored text also work during conversion, such as
\verb|\textcolor{blue}{Blue Text}|  \textcolor{blue}{Blue Text}.

### 2023-09-06 supplement/table-env/table-example.tex
pandoc citation
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

\section{Multicolumn tables}
\begin{table*}[htbp]
\begin{center}
\begin{tabular}{l | llll }
 \hline
 \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
\multicolumn{2}{c}{$Y$} \\
 \hline
  & 1 & 2 & 1 & 2 \\
 EX1  & X11 & X12 &  Y11  & Y12 \\
 EX2  & X21 & X22 &  Y21  & Y22 \\
 EX3  & X31 & X32 &  Y31  & Y32 \\
 EX4  & X41 & X42 &  Y41  & Y42\\
 EX5  & X51 & X52 &  Y51  & Y52 \\
\hline
\end{tabular}
\caption{An example table using \code{multicolumn}}
\label{table:2}
\end{center}
\end{table*}

Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.

Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
-- ToDo : include more details about MultiRow and tablenotes
\section{Complex tables}
A complex table with various other elements like figures, math, code and so on, are also supported
by \CRANpkg{texor}.  
\begin{table}[htbp]
\centering
\begin{tabular}{|l | lll |}
 \hline
 Inline Format & LaTeX Support & Web Support &  Rendering \\
 \hline
 Text       & Yes & Yes  & \emph{Hello} \\ \hline
 Image       & Yes & Yes & \begin{minipage}{0.25\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{Rlogo-5.png}
    \end{minipage} \\ \hline
 CodeBlock      & Yes & Yes & \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
\end{example}
    \end{minipage} \\ \hline
 Math      & Yes & Yes & $e = m c^2 $ \\ \hline
 Link      & Yes & Yes &  \href{www.google.com}{Google} \\ \hline
 Nested Table & Yes & No & NaN \\
\hline
\end{tabular}
\caption{Image format support in various markup/typesetting languages}
\label{table:3}
\end{table}

@@

@@

Limitaions of the \CRANpkg{texor} package in table handling includes:
\begin{itemize}
\item Usage of custom graphics/ characters like \verb|\circ| will not render in HTML properly,(using it in inline math might work).
\item Inclusion of code blocks might not always work and it is best avoided.
\item Currently only \verb|\\| is supported as the row end command/marker.
\item Some table commands/environments might not work as expected.
\item Nested tabular environments will not work as expected.
\item Colored backgrounds in tables are not supported yet.
\end{itemize}

@@

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Some common table environments.
\item Wide tables.
\item Some tables with \code{multicolumn} commands.
\item Environments such as \code{figure}/\code{code} in tables.
\end{itemize}

>>>>

\section{Multicolumn and Multirow tables}
\begin{table*}[htbp]
\begin{center}
\begin{tabular}{|l | llll |}
\hline
 \multicolumn{1}{|c |}{EXAMPLE} & \multicolumn{2}{c|}{$X$} &
\multicolumn{2}{c|}{$Y$} \\\hline
  & 1 & 2 & 1 & 2 \\\hline
 EX1  & X11 & X12 &  Y11  & Y12 \\\hline
 EX2  & X21 & X22 &  Y21  & Y22 \\\hline
\multirow{ 2}{*}{EX3} & X31 & X32 &  Y31  & Y32 \\
& X311 & X322 & Y311 & Y322\\\hline
 \multirow{ 2}{*}{EX4}  & X41 & X42 &  Y41  & Y42\\
 & X411 & X422 &  Y411  & Y422\\\hline
 EX5  & X51 & X52 &  Y51  & Y52 \\\hline
\end{tabular}
\caption{An example table using \code{multicolumn} and \code{multirow}}
\label{table:2}
\end{center}
\end{table*}

Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} and \code{multirow} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.

Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.

\section{Complex tables}
A complex table with various other elements like figures, math, code and so on, are also supported
by \CRANpkg{texor}.  
\begin{table}[htbp]
\centering
\begin{tabular}{|l | lll |}
 \hline
 Inline Format & LaTeX Support & Web Support &  Rendering \\
 \hline
 Text       & Yes & Yes  & \emph{Hello} \\ \hline
 Image       & Yes & Yes & \begin{minipage}{0.25\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{Rlogo-5.png}
    \end{minipage} \\ \hline
 CodeBlock      & Yes & Yes & \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
\end{example}
    \end{minipage} \\ \hline
 Math      & Yes & Yes & $e = m c^2 $ \\ \hline
 Link      & Yes & Yes &  \href{www.google.com}{Google} \\ \hline
 Nested Table & Yes & No & NaN \\
\hline
\end{tabular}
\caption{Image format support in various markup/typesetting languages}
\label{table:3}
\end{table}

@@

\section{Long Tables}

Pandoc supports long tables from \pkg{longtable} CTAN package as well, here is an example as table \ref{table:6}.
\begin{longtable}{|p{2cm}|p{2.5cm}|p{1cm}|p{2cm}|p{2cm}|p{2cm}|p{1cm}|}
\caption{Table of Car data from mtcars dataset}
\label{table:6}\\
\hline
\textbf{Manafacturer} & \textbf{Model and Make} & \textbf{MPG} & \textbf{Engine Cylinders} & \textbf{Engine Displacement (cu.in.)} & \textbf{Weight (1000 lbs)} & \textbf{Number of forward gears} \\
\hline
Mazda &RX4         &  21.0  & 6 & 160.0 & 2.620 &   4\\ \hline
Mazda &RX4 Wag      & 21.0  & 6 & 160.0 & 2.875  &  4\\ \hline
Datsun &710         & 22.8  & 4 &108.0 &2.320   & 4\\ \hline
Hornet &4 Drive     & 21.4  & 6 &258.0 &3.215   & 3\\ \hline
Hornet &Sportabout  & 18.7  & 8 &360.0 &3.440   & 3\\ \hline
Valiant     &      &  18.1  & 6 &225.0 &3.460   & 3\\ \hline
Duster &360        &  14.3  & 8 &360.0 &3.570   & 3\\ \hline
Merc &240D         &  24.4  & 4 &146.7 &3.190   & 4\\ \hline
Merc &230         &   22.8  & 4 &140.8 &3.150   & 4\\ \hline
Merc &280          &  19.2  & 6 &167.6 &3.440   & 4\\ \hline
Merc &280C        &   17.8  & 6 &167.6 &3.440   & 4\\ \hline
Merc &450SE        &  16.4  & 8 &275.8 &4.070   & 3\\ \hline
Merc &450SL        &  17.3  & 8 &275.8 &3.730   & 3\\ \hline
Merc &450SLC       &  15.2  & 8 &275.8 &3.780   & 3\\ \hline
Cadillac &Fleetwood & 10.4  & 8 &472.0 &5.250   & 3\\ \hline
Lincoln &Continental& 10.4  & 8 &460.0 &5.424   & 3\\ \hline
Chrysler &Imperial  & 14.7  & 8 &440.0 &5.345   & 3\\ \hline
Fiat & 128          &  32.4 &  4 & 78.7 &2.200  &  4\\ \hline
Honda &Civic        & 30.4  & 4 & 75.7 &1.615   & 4\\ \hline
Toyota &Corolla     & 33.9  & 4 & 71.1 &1.835   & 4\\ \hline
Toyota &Corona      & 21.5  & 4 &120.1 &2.465   & 3\\ \hline
Dodge &Challenger   & 15.5  & 8 &318.0 &3.520   & 3\\ \hline
AMC &Javelin        & 15.2  & 8 &304.0 &3.435   & 3\\ \hline
Cheverolet &Camaro Z28 & 13.3 & 8& 350.0 &3.840   & 3\\ \hline
Pontiac &Firebird   & 19.2  & 8 &400.0 &3.845   & 3\\ \hline
Fiat &X1-9          & 27.3  & 4 & 79.0 &1.935   & 4\\ \hline
Porsche& 914-2      & 26.0  & 4 &120.3 &2.140   & 5\\ \hline
Lotus& Europa       & 30.4  & 4 & 95.1 &1.513   & 5\\ \hline
Ford &Pantera L     & 15.8  & 8 &351.0 73.170   & 5\\ \hline
Ferrari &Dino       & 19.7  & 6 &145.0 &2.770   & 5\\ \hline
Maserati& Bora      & 15.0  & 8 &301.0 &3.570   & 5\\ \hline
Volvo &142E         & 21.4  & 4 &121.0 &2.780   & 4\\ 
\hline
\end{longtable}

@@

Limitations of the \CRANpkg{texor} package in table handling includes:
\begin{itemize}
\item Usage of custom graphics/ characters like \verb|\circ| will not render in HTML properly,(using it in inline math might work).
\item Inclusion of code blocks might not always work and it is best avoided.
\item Currently only \verb|\\| is supported as the row end command/marker.
\item Some table commands/environments might not work as expected.
\item Nested tabular environments will not work as expected.
\item Colored backgrounds in tables are not supported yet.
\end{itemize}

@@

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Some common table environments.
\item Long and Wide Tables.
\item Some tables with \code{multicolumn} and \code{multirow} commands.
\item Environments such as \code{figure}/\code{code} in tables.
\end{itemize}

\section{Acknowledgements}

Special thanks to Albert Krewinkel for his contributions towards earlier versions of Lua filters to add table numbering.
